The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-12-13 analysis/paper/paper.qmd
Apply various changes in paper

---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - Claudiu Forgaci:
      correspondence: "yes"
      email: C.Forgaci@tudelft.nl
      orcid: 0000-0003-3218-5102
      institute: tud
institute:
  - tud:
      name: Delft University of Technology
      address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format: pdf
  # docx:
  #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
  # pdf:
  #   template: "../templates/plos_latex_template.tex"
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft in 2021,
  this paper aims to describe the legacy of the EMU. To that end, it reports on the
  analysis of the graduation theses produced over 2007-2021 when the program was running.
  The text mining approach allows for a close scrutiny of the latent thematic patterns, 
  and evolutions thereof, found in the full texts of the theses. Results reveal # topics
  that are to a large extent consistent with the set-up of the program in two thematic 
  semesters at TU Delft and three possible exchange semester locations, each characterised 
  by a specific approach. The paper discusses the relevance of the findings for the
  state of the profession and education in the field. Set up as a reproducible research
  compendium accompanied by a published FAIR dataset of the theses and an interactive
  dashboard, the analysis presented can be both reproduced and applied to other datasets. 
keywords: |
  urbanism; education; post-master; text mining; topic modeling
highlights: |
  These are the highlights. 
---

@@

# TODO remove this section after all analysis has been translated into functions
# Load packages ----
## List packages to be loaded
packages <- c(
  "here",
  "tidyverse", "widyr", "readxl",   # Data manipulation and visualisation
  "furrr",                          # Parallel use of purrr::map() functions
  "tidytext", "SnowballC", "tm",    # Text processing and analysis
  "topicmodels", "LDAvis", "tsne",  # Topic modeling
  "tidygeocoder",                   # Working with spatial data
  "igraph", "ggraph", "DiagrammeR"  # Network analysis and visualization
)

@@

word_counts <- vector(mode = "numeric", length = nrow(emu))
for (i in 1:nrow(emu)) {
  word_counts[i] <- 
    tidytext::unnest_tokens(tbl = emu[i,], output = word, input = text_clean) |> nrow()
}

@@

The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft. The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005. The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field. With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation. All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism. Although located in Europe and named "European", its geographic scope was global, especially at the stage of graduation, considering that a large majority of students were of non-European origin.

In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism. Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and to some extent MSc Landscape Architecture tracks. There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit. Understanding what made the program impactful would be beneficial for targeted urbanism education and practice in response to current societal challenges. Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight into what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.

To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program. To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an average of `r round(mean(word_counts, na.rm = TRUE), -3)` words produced over the years for the duration of the program between `r paste0(min(emu$graduation_year, na.rm = TRUE) - 2, "-", max(emu$graduation_year, na.rm = TRUE))`.

### Topic modeling {#topmod}

**The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012]. At the same time, urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data. This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then. Patterns revealed in past education can potentially better inform future decisions in education in the field.

Although the corpus analysed in this study is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.

Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021]. Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a *soft clustering algorithm* in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a]. LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively. Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022a]. Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].

In addition to describing the legacy of the EMU, **this paper showcases a reproducible research workflow**. Together with a published FAIR dataset, it forms a reproducible research compendium. The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023]. The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository \[ADD CITATION\] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.

-   Give an overview of th predominant topics and their evolution over the years of the EMU program.

@@

mermaid("
graph LR
    A[Data collection]-->B[Pre-processing]
    B-->C[Information extraction]
    C-->B
    C-->D[Visualisation]
    D-->E[Results]
")
```

@@

**The dataset consists of full texts and metadata of a subset of EMU theses written in English.** The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined. The full-text documents were obtained either from the TU Delft Education Repository upon request (an API for downloading full-text data was not available at the time of data collection), or from the authors via email if not available on the repository. Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository. Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published. Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually. As a result, of the total of `r sum(theses_all_per_year$n_total)` theses presented in the EMU program, full-text documents were obtained for `r sum(theses_all_per_year$n_pdfs)` of which permissions were granted for `r nrow(emu)`. Only PDFs for which permission was obtained were included in the corpus. The final corpus contains at least one thesis for each year of graduation between `r min(emu$graduation_year)`-`r max(emu$graduation_year)` as shown in @tbl-data-selection. @fig-map-theses shows the geographic spread of the theses included in the analysis.

@@

## Information extraction

@@

Information extraction as a process of extracting information from unstructured data.

@@

<!-- [Add to vignette:] Word-level tokenisation was applied with the `unnest_tokens()` function of the `tidytext` package. This was preferred over other tokenisers as it produced results that work seamlessly with other `tidyverse` tools used in the analysis. `unnest_tokens()` removes punctuation, converts to lower-case, and separate all words into individual tokens. -->

@@

S**top words, i.e., common words or words that do not add much meaning, were removed**. Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated. As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words. <!-- We tried to create a custom global stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words, but no significant improvement was observed in the result. --> We manually constructed a list of subject-specific stop words based on our domain knowledge. Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well. Finally, titles and author names were removed as document stop words. In the next pre-processing step, we normalised the resulting words, that is, reduced them to their canonical dictionary forms, the resulting words were lemmatised [@hvitfeldt2021a]. Lemmatisation was preferred over stemming as the former tends to be more meaningful and less degrading than the latter.

```{r eval=FALSE}
ent1 <- spacyr::entity_extract(spacyr::spacy_parse(emu$text_clean[1:13]))

ent1 |> 
  filter(entity_type == "LOC") |> 
  pull(entity) |> 
  unique()

ent2 <- ent1 |> 
  mutate(entity = tolower(entity)) |> 
  group_by(entity) |> 
  mutate(bign = length(unique(doc_id))) |> 
  ungroup() |> 
  count(doc_id, entity, bign) |> 
  filter(bign > 3, nchar(entity) > 4)

dtm <- tidytext::cast_sparse(data = ent2,
                             row = doc_id,
                             column = entity,
                             value = n)

lda <- text2vec::LDA$new(n_topics = 50) 
fit <- lda$fit_transform(dtm, progressbar = F)
tm_summary <- # incomplete code

```{r eval=FALSE}
# Find locations
entity::location_entity(head_text(emu$text_clean[15], sub_end = 50000))
```

@@

Similar to clustering on numeric data, topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017]. We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003]. This means that each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.

@@

A key tuning parameter in an LDA model is the number of clusters K. A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics. Hence, finding a suitable value entails a trade-off between interpretability and specificity. This is very much dependent on the data at hand and the research question. Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach:

- TU Delft, Fall semester: Urban region networks
- TU Delft, Spring semester: Constructing the sustainable city
- IUAV Venice, exchange semester:
- KU Leuven, exchange semester:
- UPC Barcelona, exchange semester:

@@

Get top 20 words in each topic to understand what it is about

```{r lda-k-5}
#| message: false

## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
                 control = list(seed = 2023, iter = 500, verbose = 25))

@@

```{r fig-topics-per-doc}
#| fig-cap: Topic probability distribution per document

# TODO Replace this with a table containing topics, documents, and theta
#      sorted by theta in decreasing order
# What is the dominant topic in each document?
N = nrow(theta_5)
topic_prop_examples <- theta_5[1:N,]
colnames(topic_prop_examples) <- topic_names_5

vizDataFrame <- pivot_longer(
  data = data.frame(topic_prop_examples, 
                    document = factor(str_sub(rownames(topic_prop_examples), 1, 20))),
  cols = -document,
  names_to = "topic",
  values_to = "value") |> 
  group_by(document) |> 
  mutate(value_max = max(value),
         topic_max = topic[value == value_max])
ggplot(data = vizDataFrame, aes(topic, value, fill = topic_max), ylab = "proportion") + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ as.numeric(document), ncol = 10) +
  theme(legend.position = "bottom") +
  scale_fill_viridis_d(name = "Topic") +
  labs(title = "Theses as distributions over topics",
       subtitle = "Charts colored by the highest probability")

@@

```{r topics-per-corpus}

# What are the most probable topics in the entire collection?
topic_proportions <- get_topic_proportions(emu_lda_5)
# paste0(round(topic_proportions$proportion, 2), ": ", topic_proportions$topic)

# How many documents are with the highest probability in each topic?
counts_of_primary_topics <- count_docs_per_topic(emu_lda_5)
# paste0(counts_of_primary_topics$count, ": ", topic_proportions$topic)

topic_proportions |> 
  left_join(counts_of_primary_topics, by = "topic") |> 
  rename(Topic = topic,
         `Proportion per corpus` = proportion,
         `No. of documents with primary topic` = count)
```

```{r}
#| eval: false

topic_to_filter <- 1  # set manually
# or have it selected by a term in the topic name (e.g. 'airport')
topic_to_filter <- grep('water', topic_names_5)[1] 
topic_threshold <- 0.2
selected_document_indices <- which(theta_5[, topic_to_filter] >= topic_threshold)
filtered_corpus <- emu[selected_document_indices,]
# show length of filtered corpus
nrow(filtered_corpus)
```

```{r}
#| eval: false

N = nrow(emu)
topic_prop_per_year <- theta_5[1:N,]
colnames(topic_prop_per_year) <- topic_names_5

viz_df_with_year <-
  pivot_longer(
    data = data.frame(topic_prop_examples,
                      document = factor(str_sub(
                        rownames(topic_prop_examples), 1, 20))),
    cols = -document,
    names_to = "topic",
    values_to = "value") |> 
  left_join(mutate(emu, ID = as.factor(ID), "year" = grad_year), by = c("document" = "ID"))

# TODO change to discrete colors
# TODO try representing this as a line chart
# Plot topic proportions per year
viz_df_with_year |> 
  ggplot(aes(x = factor(year), y = value, fill = topic)) + 
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_viridis_d(name = "Topics") +
  xlab("Graduation year") +
  ylab("Proportion of topics") +
  labs(title = "Topic proportions per year")
```

@@

visualise_lda(phi = beta_5, 
              theta = theta_5, 
              dtm = emu_dtm)

@@

```{r topic-model-beta}

# Extract topics as mixtures of terms from LDA model
emu_topics <- tidy(emu_lda_5, matrix = "beta")

# The top n most common words in each topic
emu_top_terms <- get_top_words_per_topic(data = emu_topics, top_n = 20)

# TODO create visualisation function
emu_top_terms |> 
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip()

# TODO can this be generalised to more than two topics?

beta_spread <- emu_topics |> 
  mutate(topic = paste0("topic", topic)) |> 
  spread(topic, beta) |> 
  filter(topic1 > 0.001, topic2 > .001) |> 
  mutate(log_ratio = log2(topic2/topic1))

beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  slice_max(abs(log_ratio), n = 20) %>% 
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(log_ratio, term)) +
  geom_col() +
  labs(x = "Log2 ratio of beta in topic 2 / topic 1", y = NULL)
```

# TODO check if this plot is useful and, if not, remove it
emu_documents |> 
  inner_join(emu |> 
               select(ID, grad_year), 
             by = c("document" = "ID")) |> 
  ggplot(aes(x = factor(gradyear), y = topic, fill = factor(topic))) +
  geom_col()
```

## Overall patterns

The tf-idf statistic shows that...

Topics such as ... have gained importance over the years, while ... have lost prominence.

```{r top-words-document}
#| eval: false

# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_words, id_col = "ID", top_n = 10)
```

```{r top-words-corpus}

# Which are the top n most frequently used words in all theses?
terms <- as.data.frame(posterior(emu_lda_5)$terms)
rownames(terms) <- topic_names_5
terms <- terms |> 
  mutate(topic = rownames(terms)) |> 
  pivot_longer(-topic,
               names_to = "term",
               values_to = "prob") |> 
  group_by(term) |> 
  mutate(max_topic = topic[which.max(prob)]) |> 
  filter(topic == max_topic) |> 
  ungroup()

emu_words_topics <- emu_words |> 
  left_join(terms, by = c("word" = "term"))

n = 20
get_top_words_per_corpus(data = emu_words, top_n = n) |>
  left_join(terms, by = c("word" = "term")) |> 
  ggplot(aes(reorder(word, n), n)) +
  geom_col(aes(fill = max_topic)) +
  geom_text(aes(label = n), size = 2, hjust = 1.1) +
  coord_flip() +
  xlab("Word") +
  ylab("Frequency") +
  labs(title = paste0("Top ", n, " most used words in the corpus of theses")) +
  theme_minimal() +
  theme(panel.grid = element_blank())
```

A frequency count of ngrams reveals collocations like "public space" or "socio-spatial segregation" across the corpus. To include these in the topic model, we concatenated them with an underscore, so that they are considered one word during the word-level tokenisation. Only bigrams and trigrams, that is two- and three-word collocations, were extracted from the corpus.

```{r bigram-tfidf}

emu_bigrams <- emu |> 
  get_ngrams(n = 2, id_col = "ID", text_col = "text_clean", stem = TRUE)

bigram_counts <- emu_bigrams |> 
  count(w_1, w_2, sort = TRUE)

bigram_tf_idf <- emu_bigrams |> 
  count(ID, ngram) |> 
  bind_tf_idf(ngram, ID, n) |> 
  arrange(desc(tf_idf))

bigram_tf_idf
```

```{r trigram-tfidf}

emu_trigrams <- emu |> 
  get_ngrams(n = 3, id_col = "ID", text_col = "text_raw")

trigram_counts <- emu_trigrams |> 
  count(w_1, w_2, w_3, sort = TRUE)

trigram_tf_idf <- emu_trigrams |> 
  count(ID, ngram) |> 
  bind_tf_idf(ngram, ID, n) |> 
  arrange(desc(tf_idf))

trigram_tf_idf

trigram_tf_idf |> arrange(desc(n))
```

```{r}
## Show network of bigrams
bigram_graph <- bigram_counts |> 
  filter(n > 100) |> 
  graph_from_data_frame()

bigram_graph

set.seed(2023)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

ggraph(bigram_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a) +
  geom_node_point(color = "lightblue", linewidth = 5) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1) +
  theme_void()

## Higgh frequency bigrams ----

emu_bigrams_top <- emu_bigrams |>
  group_by(ID) |>
  count(ngram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  ungroup() |>
  mutate(bigram = reorder(ngram, n))

emu_bigrams_top_all <- emu_bigrams |>
  count(ngram, sort = TRUE) |>
  slice_max(n, n = 20) |>        ### top_n is superseded; better use slice_max instead
  # filter(n > 50)                ### filter words with a minimum count
  mutate(ngram = reorder(ngram, n))

emu_bigrams_top |>
  filter(ID == unique(ID)[1:20]) |> 
  ggplot(aes(x = ngram, y = n, fill = ID)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ ID, scales = "free", ncol = 2) +
  theme(legend.position = "none") 
```

@@

>>>>

---
title: "The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach"
author:
  - Claudiu Forgaci:
      correspondence: "yes"
      email: C.Forgaci@tudelft.nl
      orcid: 0000-0003-3218-5102
      institute: tud
institute:
  - tud:
      name: Delft University of Technology
      address: Julianalaan 134, 2628 BL, Delft, Zuid-Holland, The Netherlands
title-block-published: "Last updated"  
date: now
date-format: long
format: pdf
  # docx:
  #   reference-doc: "../templates/template.docx" # Insert path for the DOCX file
  # pdf:
  #   template: "../templates/plos_latex_template.tex"
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  This paper aims to describe the legacy of the European post-Master of Urbanism 
  (EMU) that ran between 2005 and 2021 at TU Delft. To that end, it reports on the
  analysis of the graduation theses produced over the duration of the program.
  The text mining approach allows for a close scrutiny of the latent thematic patterns, 
  and evolutions thereof, found in the full texts of the theses. Results reveal 5 topics
  that are to a large extent consistent with the set-up of the program in two thematic 
  semesters at TU Delft and three possible exchange semester locations, each characterised 
  by a specific approach. The paper discusses the relevance of the findings for the
  state of the profession and education in the field. Set up as a research
  compendium accompanied by a published FAIR dataset and an interactive dashboard, 
  the research is fully reproudcible and enables further analysis. 
keywords: |
  urbanism; education; post-master; text mining; topic modeling
highlights: |
  These are the highlights. 
---

@@

# TODO remove unused packages after all analysis has been translated into functions
# Load packages ----
## List packages to be loaded
packages <- c(
  "here",
  "tidyverse", "widyr", "readxl",   # Data manipulation and visualisation
  "furrr",                          # Parallel use of purrr::map() functions
  "tidytext", "SnowballC", "tm",    # Text processing and analysis
  "topicmodels", "LDAvis", "tsne",  # Topic modeling
  "tidygeocoder",                   # Working with spatial data
  "igraph", "ggraph", "DiagrammeR"  # Network analysis and visualization
)

@@

word_counts <- get_word_counts(emu, text_col = "text_clean")

@@

The research presented in this paper was prompted by the closure of the European post-Master of Urbanism (EMU) at TU Delft. The EMU is an advanced program started jointly by TU Delft, KU Leuven, UPC Barcelona and Università IUAV di Venezia in 2005. The program distinguished itself internationally as a post-master open only to applicants with a prior MSc degree and practical experience in urbanism or a related field. With a combined research and design curriculum, it provided a bridge towards both advanced practice and research trajectories after graduation. All students followed an exchange semester at one of the other three program universities, allowing them to have a broad, international view on the field of urbanism. Although located in Europe and named "European", its geographic scope was global, visible especially at the stage of graduation, considering that a large majority of students were of non-European origin and chose to work on a case from their home location.

In 2021, TU Delft stepped out from the program, but its approach has remained imprinted on the TU Delft approach to urbanism. Its core topics Urban Region Networks (Semester 1) and Constructing Sustainable Urban Landscapes (Semester 2), as well as the influence of the partner universities, mostly IUAV Venice where most TU-Delft-hosted students spent their exchange semesters, are visible in the current MSc Urbanism and, to some extent, MSc Landscape Architecture tracks. There is general agreement on the considerable impact of the EMU on education and practice targeting sustainable urbanism, but why and how it is so remains implicit. Understanding what made the program impactful would be beneficial for targeted urbanism education and practice in response to current societal challenges. Having a close look at the assignments that were addressed by the program as well as research and design approaches used to tackle those topics throughout the years can provide valuable insight into what are the core, stable elements of urbanism practice, on the one hand, and how does the field need to adapt to changing societal challenges, on the other.

To describe the legacy of the EMU program, including the distinctive features of its didactic approach, this paper aims to reveal the main topics taught in it and how those topics had evolved over the years of the program. To that end, the research presented in this paper employed a text mining approach targeting the output of the EMU: 96 theses with an rounded average of `r round(mean(word_counts, na.rm = TRUE), -3) |> format()` words produced over the years for the duration of the program between `r paste0(min(emu$grad_year, na.rm = TRUE) - 2, "-", max(emu$grad_year, na.rm = TRUE))`. Although the corpus of thesis is relatively small, it is large enough to make a manually conducted quantitative analytical approach unfeasible.

## A text mining approach {#textmin}

**The rise of computational social sciences** [@wallach2018] provides tools and methods to summarise large sets of data that would not be possible with human annotation [@blei2012]. Often such data is available in unstructured textual form and formats that are not easily machine-readable, such as Adobe PDF or Microsoft Word documents [@benchimol2022]. While quantitative text analysis have been widely adopted in computationla social sciences such as linguistics, communication, and political science [@benchimol2022], urbanism research, in general, and urbanism education studies, in particular, rarely use statistical methods such as topic modeling to analyse unstructured data. This is especially a missed opportunity considering that theses are representative outputs of educational programs that are increasingly available in digital repositories - the TU Delft repository was established in 2007, which is also the first graduation year of the EMU, and has been steadily growing since then. Patterns revealed in past education can potentially better inform future decisions in education in the field.

Acknowledging that opportunity, **this paper leverages topic modeling**, a branch of Natural Language Processing, as an unsupervised machine learning method to reveal latent semantic structure and relationships between latent topics in a large document collection [@muchene2021]. Latent Dirichlet Allocation (LDA) is broadly adopted in topic modeling [@muchene2021] as it uses a soft clustering algorithm in which objects can be part of multiple clusters (mixed membership approach) as opposed to hard clustering algorithms (single membership approaches) [@rüdiger2022a]. LDA uses as input a bag-of-words (BOW) model, more generally called the vector space model (VSM), in which documents in a corpus are decomposed into two low-dimensional matrices for document-topic distribution and word-topic distribution, respectively. Although topic modeling in general automates the process of identifying topics through clustering, those topics need to be manually interpreted and labelled after the application of the algorithm [@rüdiger2022a]. Moreover, arguing for the right algorithm and a suitable number of clusters remains a challenge [@rüdiger2022].

In addition to describing the legacy of the EMU, **this paper showcases a reproducible research workflow**. Together with a published FAIR dataset, it forms a reproducible research compendium. The analyses presented in this paper were carried out in the free and open-source statistical programming software R [@rcoreteam2023]. The research compendium is available at \[ADD CITATION\], the data is published in th 4TU.ResearchData repository [@forgaci2023] and it is available for exploration in a Shiny dashboard at <http://mintemu_shiny.bk.tudelft.nl:3838/mintemu/>.

@@

mermaid("graph LR
    A[Data collection]-->B[Pre-processing]
    B-->C[Information extraction]
    C-->B
    C-->D[Visualisation]
    D-->E[Results]
    ")
```

@@

**The dataset consists of full texts and metadata of a subset of EMU theses written in English.** The full texts are available as PDF documents with unstructured layouts typical to an urbanism project in which text and various types of graphics are combined <!--add figure-->. The full-text documents were obtained either from the TU Delft Education Repository upon request (an API for downloading full-text data was not available at the time of data collection), or from the authors via email if not available on the repository. Metadata was downloaded from the TU Delft Education Repository or obtained from the EMU program coordination records when unavailable on the repository. Even though repository entries are openly accessible, licenses were not specified and thus it was unclear if the content of the theses can be mined and if the resulting data can be published. Considering the relatively small size of the corpus and the possibility of directly contacting the alumni network, permissions were requested individually. As a result, of the total of `r sum(theses_all_per_year$n_total)` theses presented in the EMU program, full-text documents were obtained for `r sum(theses_all_per_year$n_pdfs)` of which permissions were granted for `r nrow(emu)`. Only PDFs for which permission was obtained were included in the corpus. The final corpus contains at least one thesis for each year of graduation between `r min(emu$grad_year)`-`r max(emu$grad_year)` as shown in @tbl-data-selection. @fig-map-theses shows the geographic spread of the theses included in the analysis.

@@

## Information extraction {#infoext}

@@

Information extraction is a process of extracting information from unstructured data following the steps of tokenisation, construction of a Document-Term Matrix (DTM), fitting the topic model

@@

@@

**Stop words, i.e., common words or words that do not add much meaning, were removed**. Global, subject-specific and document stop words [@hvitfeldt2021a] were differentiated. As all theses were written in English, the pre-made "snowball" lexicon was used to remove global stop words. We manually constructed a list of subject-specific stop words based on our domain knowledge. Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as subject-specific stop words as well. Finally, titles and author names were removed as document stop words. In the next pre-processing step, we normalised the resulting words, that is, reduced them to their canonical dictionary forms, the resulting words were lemmatised [@hvitfeldt2021a]. Lemmatisation was preferred over stemming as the former tends to be more meaningful and less degrading than the latter.

@@

Similar to clustering on numeric data [@rüdiger2022], topic modelling is a method of unsupervised classification of topics found in a collection of documents [@silge2017]. We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. LDA is a Bayesian generative topic model which assumes that each topic is a distribution over words and each document is a distribution over topics [@blei2012; @grün2011; @blei2003]. This means that each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\theta$ (theta) statistic.

@@

A key tuning parameter (i.e., a manually specified parameter) in an LDA model is the number of clusters K. A K with a smaller value will yield more general topics whereas a K with a high value will result in more specific topics. Hence, finding a suitable value entails a trade-off between interpretability and specificity. This is very much dependent on the data at hand and the research question. Given the small size of the corpus and the way the program was structured, the value of K = 5 was determined qualitatively considering the input received by the students in the two thematic semesters at TU Delft and the three alternative exchange semesters, each with a very distinct approach.

| University    | Semester        | Topic                                            |
|------------------|------------------|------------------------------------|
| TU Delft      | Fall semester   | Urban region networks                            |
|               | Spring semester | Constructing the sustainable city                |
| IUAV Venice   | Fall semester   | Territories of dispersion: Situations, Scenarios |
| KU Leuven     | Fall semester   |                                                  |
| UPC Barcelona | Fall semester   | Urban Transformations                            |

Each topic is assigned a pseudo-name comprising the top 5 most frequent words of the topic. The top 20 words were used to interpret and describe the meaning of the topics. Overall results per topic are visualised with bar charts and the evolution of topics and top words is visualised with line charts.

Using the $\theta$ statistic, the distribution of documents over topics is visualised to classify documents accoriding to their dominant topic and to show to what extent a document is specific to a topic or spread across topics.

@@

```{r lda}
#| include: false

## K = 5 considering the 2 TUD semesters and the 3 exchange semesters combined 
emu_lda_5 <- LDA(emu_dtm, k = 5, method="Gibbs",
                 control = list(seed = 2023, iter = 500, verbose = 100))

@@

<!-- Add intro paragraph -->

```{r fig-topics-doc}
#| fig-cap: Topic probability distribution per document

# TODO Replace this with a table containing topics, documents, and theta
#      sorted by theta in decreasing order
# What is the dominant topic in each document?
topic_prop <- as.data.frame(theta_5)
colnames(topic_prop) <- paste0(paste0("Topic ", 1:5, ": "), topic_names_5)

topic_prop_viz <- 
  pivot_longer(data = data.frame(topic_prop,
                                 document = factor(rownames(topic_prop)),
                                 check.names = FALSE),
    cols = -document,
    names_to = "topic",
    values_to = "value") |> 
  group_by(document) |> 
  mutate(value_max = max(value),
         topic_max = topic[value == value_max])

p <- ggplot(data = topic_prop_viz, 
       aes(topic, value, fill = topic_max), 
       ylab = "proportion") + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  labs(title = "Theses as distributions over topics",
       subtitle = "Charts colored by the topic with the highest probability") +
  ylab("Probability") +
  scale_x_discrete(labels = paste0("Topic ", 1:5)) +
  coord_flip() +
  facet_wrap(~ factor(document, levels = rownames(topic_prop)), ncol = 10) +
  theme(legend.position = "bottom",
        legend.justification = "left",
        legend.direction = "vertical",
        axis.title.y = element_blank()) +
  scale_fill_viridis_d(name = "Dominant topic")

p_tab <- ggplotGrob(p)
# print(p_tab)

gtable_filter_remove <- function (x, name, trim = TRUE){
    matches <- !(x$layout$name %in% name)
    x$layout <- x$layout[matches, , drop = FALSE]
    x$grobs <- x$grobs[matches]
    if (trim) 
        x <- gtable_trim(x)
    x
}

select_grobs <- function() {
  x <- vector()
  for (i in 2:10) {
    for (j in 1:10) {
      x <- c(x, paste0("axis-l-", i, "-", j))
      x <- c(x, paste0("axis-b-", i, "-", j))
    }
  }
  x
}

p_filtered <- gtable_filter_remove(p_tab, name = select_grobs(), trim = FALSE)

library(grid)
grid::grid.newpage()
grid::grid.draw(p_filtered)

@@

```{r topic-proportions-corpus}

# What is the mean probability of the documents being part of each topic?
topic_proportions <- get_topic_proportions(emu_lda_5)

# How many documents are with the highest probability in each topic?
counts_of_primary_topics <- count_docs_per_topic(emu_lda_5)

# Show summary table of mean probabilities and document counts per topic
topic_proportions |> 
  left_join(counts_of_primary_topics, by = "topic") |> 
  rename(Topic = topic,
         `Proportion per corpus` = proportion,
         `No. of documents with primary topic` = count)
```

The top 20 most frequently used words in the corpus (@fig-top-words-corpus), do not reflect the mean proportion ...

```{r fig-top-words-corpus}
# Which are the top n most frequently used words in all theses?
vis_top_words_per_corpus(lda = emu_lda_5, words = emu_words, top_n = 20)
```

```{r top-words-document}
# Which are the top n most frequently used words in each thesis?
get_top_words_per_document(data = emu_words, id_col = "ID", top_n = 3)
```

```{r top-words-topic}
vis_top_words_per_topic(data = emu_lda_5, top_n = 20)
```

@@

vis_lda(phi = beta_5,
        theta = theta_5,
        dtm = emu_dtm)

@@

@@

### Topic 1: `r topic_names_5[1]`

### Topic 2: `r topic_names_5[2]`

### Topic 3: `r topic_names_5[3]`

### Topic 4: `r topic_names_5[4]`

### Topic 5: `r topic_names_5[5]`

```{r fig-topics-evolution}
#| eval: false

topic_prop_per_year <- theta_5[1:nrow(emu),]
colnames(topic_prop_per_year) <- topic_names_5
